---
title: "Power_analysis_FEFA"
author: "Maria Lee"
date: "9/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pwr)
library(ggplot2)

```

Power analysis in order to figure out how many statistical tests we can do in the FEFA KASP study (given that data has already been collected). 
In the first part, we want to compare performance on FEFA between two groups - the patients and the controls with regards to different emotions. 
In the second part we wish to correlate performance on FEFA to symptom levels. Is there a correlation between level of psychotic symptoms and inability to identify emotions?  

We have 69 patients with FEFA data and 52 control subjects (these numbers were used as the power calculation was performed before scrutinizing subjects in great detail. 2 patients and 1 control were later excluded due to drug use and being a statistical outlier)

# Part 1, group comparison 

What effect size do we expect in the group comparison? 
-0.27 to -1.35 for the whole test (general emotion detection), based on previous literature (for example Miklovic, 2020)


```{r}
n1 <- 69 #patients (this is the n to be used for correlation calculations as well)
n2 <- 52 #controls
alpha_level <- seq(from = 0.0001, to = 0.05,by = 0.0005)

#declare empty vectors as placeholders
powercurve_2 <- c() 
powercurve_5_ <- c()
powercurve_8 <- c()

#for loop to create all powercurves#
for (s in 1:length(alpha_level)) {
  output_2 <- pwr.t2n.test(n = n1, n2 = n2, d = 0.2, sig.level = alpha_level[s], alternative = "greater") 
  powercurve_2[s] <- output_2$power #use dollar sign to fetch only power "value", and assign it to "powercurve_2"-vector
  
  output_5_ <- pwr.t2n.test(n = n1, n2 = n2, d = 0.5, sig.level = alpha_level[s], alternative = "greater")
  powercurve_5_[s] <- output_5_$power
  
  output_8 <- pwr.t2n.test(n = n1, n2 = n2, d = 0.8, sig.level = alpha_level[s], alternative = "greater")
  powercurve_8[s] <- output_8$power
}

#Before plotting data or analysing data in R, ALWAYS create a dataframe

power_frame2 <- data.frame(alpha_level, powercurve_2, powercurve_5_, powercurve_8)
power_frame2$alpha_level

#Plot
plot(power_frame2$alpha_level, power_frame2$powercurve_2, ylim = c(0.00, 1.00), 
     ylab = "Power level", xlim = c(0.0001, 0.05), xlab = "Alpha level", 
     type = "o", col = "red", main = "Powercurves for d = 0.2 (red), 0.5 (black), 0.8 (green)")

par(new=TRUE)

plot(power_frame2$alpha_level, power_frame2$powercurve_5_, ylim = c(0.00, 1.00), 
     ylab = "", xlim = c(0.0001, 0.05), xlab = "", 
     type = "o", col = "black", main = "Powercurves for d = 0.2 (red), 0.5 (black), 0.8 (green)")

par(new=TRUE)

plot(power_frame2$alpha_level, power_frame2$powercurve_8, ylim = c(0.00, 1.00), 
     ylab = "", xlim = c(0.0001, 0.05), xlab = "", 
     type = "o", col = "green", main = "Powercurves for d = 0.2 (red), 0.5 (black), 0.8 (green)")


```


The graph displays curves for power levels depending on the effect size (in this case a Cohen's D of 0.2, 0.5 and 0.8) given our sample size. The x-axis displays the alpha-level which would be 0.05 for one comparison, but will need to be adjusted for multiple comparison. Our power to detect a difference between groups decreases with the number of statistical tests performed. 

Our aim was to be able to detect a group difference of a medium to large effect size approx. Cohen’s d = 0.7). Given these power curves our estimation is that we are able to perform 5 group comparisons, will retaining about 80 % power. 


# Part 2, correlation between symptoms and FEFA scores - only for patients

What effect size is reasonable to expect when correlating psychotic symptoms to FAR ability? 
In the literature, r-values ranged between -0.269 and -0.453 (Lescynzka, 2015 and Mitrovic, 2020)


```{r}

#declare empty vectors as placeholders
powercurve_3 <- c() 
powercurve_5 <- c()
powercurve_7 <- c()

#for loop to create all powercurves#
for (s in 1:length(alpha_level)) {
  output_3 <- pwr.r.test(n = n1, r = 0.4, sig.level = alpha_level[s]) 
  powercurve_3[s] <- output_3$power #use dollar sign to fetch only power "value", and assign it to "powercurve_3"-vector
  
  output_5 <- pwr.r.test(n = n1, r = 0.5, sig.level = alpha_level[s])
  powercurve_5[s] <- output_5$power
  
  output_7 <- pwr.r.test(n = n1, r = 0.7, sig.level = alpha_level[s])
  powercurve_7[s] <- output_7$power
}

#Before plotting data or analysing data in R, ALWAYS create a dataframe

power_frame <- data.frame(alpha_level, powercurve_3, powercurve_5, powercurve_7)
power_frame$alpha_level


#Plot
plot(power_frame$alpha_level, power_frame$powercurve_3, ylim = c(0.00, 1.00), 
     ylab = "Power level", xlim = c(0.0001, 0.05), xlab = "Alpha level", 
     type = "o", col = "red", main = "Powercurves for r = 0.3 (red), 0.5 (black), 0.7 (green)")

par(new=TRUE)

plot(power_frame$alpha_level, power_frame$powercurve_5, ylim = c(0.00, 1.00), 
     ylab = "", xlim = c(0.0001, 0.05), xlab = "", 
     type = "o", col = "black", main = "Powercurves for r = 0.3 (red), 0.5 (black), 0.7 (green)")

par(new=TRUE)

plot(power_frame$alpha_level, power_frame$powercurve_7, ylim = c(0.00, 1.00), 
     ylab = "", xlim = c(0.0001, 0.05), xlab = "", 
     type = "o", col = "green", main = "Powercurves for r = 0.3 (red), 0.5 (black), 0.7 (green)")



```

The graph displays curves for power levels depending on the effect size (in this case a Pearson's r of 0.3, 0.5 and 0.7) given our sample size (only the patient group). The x-axis displays the alpha-level which would be 0.05 for one comparison, but will need to be adjusted for multiple comparison. Our power to detect a significant correlation decreases with the number of statistical tests performed. 


Our aim in the second part of the study was to be able to detect a correlation of moderate magnitude (considered to be an r = 0.5) between psychotic symptom levels and FEFA test resuls. Given the power curves above we estimated that we were able to perform 6 correlations while still remaining 80% power. 


